---
title: "Evaluating the impact of lexico-grammatical tagging systems on multidimensional analysis"
author: "Andreas Blombach"
date: last-modified
date-format: long
format:
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: false
    code-tools: true
    theme: yeti
    df-print: paged
    fig-width: 8
    fig-height: 5
    link-external-icon: true
---

## Setup
```{r}
#| message: false
library(tidyverse)
library(readxl)
library(psych)
library(patchwork)
library(skimr)
library(paletteer)
library(see)
library(easystats)
library(stargazer)
library(lme4)
library(ggtext)
```

Features -- this is a dataframe containing all linguistic features from Biber
(1988), their names in the Biber tagger, in pseudobibeR/pybiber, MAT, BiberPlus,
and, if applicable, in Biberpy and MFTE. It also contains their means, minimum
and maximum values and standard deviations as well as their factor loadings in
the original study (Biber 1988).
```{r}
#| warning: false
#| message: false
feats_loadings <- read_excel("features_loadings_biber1988.xlsx")
```

Feature selection for dimensions according to Biber (1988) -- this is a
dataframe containing linguistic features and their positive (1) or negative (-1)
association with a given dimension.
```{r}
#| warning: false
#| message: false
biber_dims <- read_csv("dims_biber1988.csv")
```

Scores from Egbert & Mahlberg (2020) that we're trying to replicate:
```{r}
#| warning: false
#| message: false
full_scores_egbert_mahlberg <- read_csv("egbert_mahlberg_dimension_scores.csv")
```


### MAT features etc.
NB: Nini standardises features differently. We will deal with this later during
*z*-standardisation.
```{r}
#| warning: false
#| message: false
files_statistics <- list.files("output_MAT/", r"(Statistics.+?uotes.csv)",
                               recursive = TRUE, include.dirs = TRUE,
                               full.names = TRUE)
# files_zscores <- list.files("output_MAT/", r"(Zscores.+?uotes.csv)",
#                             recursive = TRUE, include.dirs = TRUE,
#                             full.names = TRUE)
# files_dimensions <- list.files("output_MAT/", r"(Dimensions.+?uotes.csv)",
#                                recursive = TRUE, include.dirs = TRUE,
#                                full.names = TRUE)

if(exists("nini_statistics")) rm(nini_statistics)

for (file in files_statistics) {
  corpus <- str_extract(file, r"(Statistics/Statistics_(.+?)[ _])", group = 1)
  qnq <- str_extract(file, r"([ _](Quotes|Nonquotes))", group = 1)
  if (!exists("nini_statistics")) {
    nini_statistics <- read_csv(file) |>
      select(!last_col()) |> # last column is garbage
      head(-1) |> # summary statistics are not needed
      mutate(corpus = corpus,
             subcorpus = qnq) |>
      select(corpus,
             subcorpus,
             Filename,
             everything())
  } else {
    nini_statistics <- nini_statistics |>
      bind_rows(read_csv(file) |>
                  select(!last_col()) |>
                  head(-1) |>
                  mutate(corpus = corpus,
                         subcorpus = qnq) |>
                  select(corpus,
                         subcorpus,
                         Filename,
                         everything()))
  }
}

nini_statistics <- nini_statistics |>
  mutate(subcorpus = str_to_lower(subcorpus))
```


### pseudobibeR
[pseudobibeR](https://github.com/browndw/pseudobibeR) by David West Brown is an
R package that extracts Biber's original 67 linguistic features from output from
either udpipe or spaCy. Since udpipe's output is in CoNLL-U format, it is easy
to adapt pseudobibeR to also accept CoNLL-U output from state-of-the-art systems
like Stanza. NB: Outputs from different systems may have their own peculiarities,
mostly depending on the version of the UD treebanks they were trained on. While
pseudobibeR has been specifically adapted to udpipe and spaCy output, this isn't
the case for Stanza.

#### Using Stanza
```{r}
#| warning: false
#| message: false
pseudobiber_stanza_statistics <- read_csv("output_pseudobiber/quotes_nonquotes_corpus/stanza_pseudobiber_features.csv")
```

#### Using spaCy (via `spacyr`)
```{r}
#| warning: false
#| message: false
pseudobiber_spacy_sm_statistics <- read_csv("output_pseudobiber/quotes_nonquotes_corpus/spacy_sm_pseudobiber_features.csv")
pseudobiber_spacy_md_statistics <- read_csv("output_pseudobiber/quotes_nonquotes_corpus/spacy_md_pseudobiber_features.csv")
pseudobiber_spacy_lg_statistics <- read_csv("output_pseudobiber/quotes_nonquotes_corpus/spacy_lg_pseudobiber_features.csv")
```


### pybiber
[pybiber](https://github.com/browndw/pybiber) is a reimplementation of
pseudobibeR in Python which only uses spaCy. The results *should* be the same as
those of pseudobibeR (using spaCy), but since there is no back-and-forth between
R and Python, execution should be faster.
```{r}
#| warning: false
#| message: false
pybiber_spacy_sm_statistics <- read_tsv("output_pybiber/output_pybiber_sm.csv")
pybiber_spacy_md_statistics <- read_tsv("output_pybiber/output_pybiber_md.csv")
pybiber_spacy_lg_statistics <- read_tsv("output_pybiber/output_pybiber_lg.csv")
pybiber_spacy_trf_statistics <- read_tsv("output_pybiber/output_pybiber_trf.csv")
```

```{r}
pybiber_spacy_sm_statistics <- pybiber_spacy_sm_statistics |>
  mutate(corpus = str_extract(doc_id, r"(^(\w+?)_)", group = 1),
  subcorpus = str_extract(doc_id, r"(_((non)?quotes))", group = 1),
  doc_id = str_extract(doc_id, r"(_(\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything())
pybiber_spacy_md_statistics <- pybiber_spacy_md_statistics |>
  mutate(corpus = str_extract(doc_id, r"(^(\w+?)_)", group = 1),
  subcorpus = str_extract(doc_id, r"(_((non)?quotes))", group = 1),
  doc_id = str_extract(doc_id, r"(_(\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything())
pybiber_spacy_lg_statistics <- pybiber_spacy_lg_statistics |>
  mutate(corpus = str_extract(doc_id, r"(^(\w+?)_)", group = 1),
  subcorpus = str_extract(doc_id, r"(_((non)?quotes))", group = 1),
  doc_id = str_extract(doc_id, r"(_(\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything())
pybiber_spacy_trf_statistics <- pybiber_spacy_trf_statistics |>
  mutate(corpus = str_extract(doc_id, r"(^(\w+?)_)", group = 1),
  subcorpus = str_extract(doc_id, r"(_((non)?quotes))", group = 1),
  doc_id = str_extract(doc_id, r"(_(\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything())
```

### BiberPlus
[BiberPlus](https://github.com/davidjurgens/biberplus) is another Python package
that extracts Biber's original 67 features. It also implements additional
features, "inspired by the work of Grieve, Clarke, and colleagues".

Initial part-of-speech tagging is done via spaCy's small (sm) model (with no
option to change the model). All features except for type-token ratio and
average word length are then computed and averaged over 1,000 token windows (a
notable difference to all other programs). TTR is based on the first 400 tokens
as in Biber's original implementation.

To get an idea of how much influence the window size has, we re-ran Biberplus
with 10,000 token windows. These results then have to be normalised to be
comparable.

```{r}
#| warning: false
#| message: false
biberplus_statistics <- read_csv("output_biberplus/output_biberplus.csv")
biberplus_statistics <- biberplus_statistics |>
  mutate(corpus = str_extract(doc_id, r"(^(\w+?)_)", group = 1),
  subcorpus = str_extract(doc_id, r"(_((non)?quotes))", group = 1),
  doc_id = str_extract(doc_id, r"(_(\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything())

biberplus_10k_statistics <- read_csv("output_biberplus/output_biberplus_10k.csv")
biberplus_10k_statistics <- biberplus_10k_statistics |>
  mutate(corpus = str_extract(doc_id, r"(^(\w+?)_)", group = 1),
  subcorpus = str_extract(doc_id, r"(_((non)?quotes))", group = 1),
  doc_id = str_extract(doc_id, r"(_(\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything()) |>
  mutate(across(QUAN:X, ~ .x / 10))
```

### Biberpy
[Biberpy](https://github.com/ssharoff/biberpy) is a collection of Python scripts
by Serge Sharoff designed to extract Biber's linguistic features, although only
48 features are currently implemented. It works with JSON output from spaCy (a
script to produce such output is provided) or CONLL style output from other NLP
pipelines (like UDPipe).

All scripts are supposed to be run from the command line.

We adapted Biberpy's spaCy script to be used with an entire folder of .txt
files, then ran its biber-dim.py script to extract normalised feature counts:
```
python spacydir2json.py quotes_nonquotes_corpus en_core_web_sm >quotes_nonquotes.json
python biber-dim.py -f json -l en -i quotes_nonquotes.json -o output_biberpy_sm.tsv
```

Since many features are normalised by clause count instead of document length
(= token count), we changed this behaviour to make Biberpy more comparable to
the other systems. We also need to multiply counts by 1,000 (normalised counts
per 1,000 tokens).

```{r}
#| warning: false
#| message: false
biberpy_statistics <- read_tsv("output_biberpy/output_biberpy_sm.tsv")
biberpy_statistics <- biberpy_statistics |>
  mutate(corpus = str_extract(doc_id, r"(^(\w+?)_)", group = 1),
  subcorpus = str_extract(doc_id, r"(_((non)?quotes))", group = 1),
  doc_id = str_extract(doc_id, r"(_(\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything()) |>
  mutate(across(A01.pastVerbs:last_col() & !J43.TTR:J44.wordLength, ~ .x * 1000))
```


## Computing *z*-scores
### MAT
To get comparable scores, type-token ratio has to be between 0 and 100 (which is
already unusual, but MAT's TTR values have a really weird range), and features
normalised on the total number of tokens need to be multiplied by 10.
```{r}
nini_zscores <- nini_statistics |>
  mutate(TTR = TTR / 4) |>
  mutate(across(AMP:XX0, ~ .x * 10)) |>
  pivot_longer(cols = AWL:XX0, names_to = "feature_MAT") |>
  left_join(feats_loadings |> select(mean, sd, `Linguistic feature`:feature_MAT, factor1:factor5), by = "feature_MAT") |>
  rename(doc_id = Filename) |>
  mutate(doc_id = str_extract(doc_id, r"((\w+)_(non)?quotes)", group = 1)) |>
  select(doc_id, corpus, subcorpus, everything()) |>
  mutate(zscore = (value - mean) / sd) |>
  arrange(corpus, doc_id, subcorpus, feature_pseudobiber) # |>
  # mutate(zscore = ifelse(zscore > 5, 5, zscore), # possible correction for small and large values
  #        zscore = ifelse(zscore < -5, -5, zscore))
nini_zscores
```

### pseudobibeR
Convenience function to compute *z*-scores:
```{r}
get_zscores <- function(statistics, correction = FALSE) {
  zscores <- statistics |>
    mutate(f_43_type_token = f_43_type_token * 100) |>
    pivot_longer(cols = f_01_past_tense:last_col(), names_to = "feature_pseudobiber") |>
    left_join(feats_loadings |> select(mean, sd, `Linguistic feature`:feature_MAT, factor1:factor5), by = "feature_pseudobiber") |>
    mutate(zscore = (value - mean) / sd)
  
  if (correction) {
    zscores <- zscores |>
      mutate(zscore = ifelse(zscore > 5, 5, zscore), # correction for small and large values
             zscore = ifelse(zscore < -5, -5, zscore))
  }
  
  return(zscores)
}
```


#### Using Stanza
```{r}
pseudobiber_stanza_zscores <- pseudobiber_stanza_statistics |>
  get_zscores()
pseudobiber_stanza_zscores
```

#### Using spaCy
```{r}
pseudobiber_spacy_sm_zscores <- pseudobiber_spacy_sm_statistics |>
  get_zscores()
pseudobiber_spacy_sm_zscores

pseudobiber_spacy_md_zscores <- pseudobiber_spacy_md_statistics |>
  get_zscores()
pseudobiber_spacy_md_zscores

pseudobiber_spacy_lg_zscores <- pseudobiber_spacy_lg_statistics |>
  get_zscores()
pseudobiber_spacy_lg_zscores
```

### pybiber
```{r}
pybiber_spacy_sm_zscores <- pybiber_spacy_sm_statistics |>
  get_zscores()
pybiber_spacy_sm_zscores

pybiber_spacy_md_zscores <- pybiber_spacy_md_statistics |>
  get_zscores()
pybiber_spacy_md_zscores

pybiber_spacy_lg_zscores <- pybiber_spacy_lg_statistics |>
  get_zscores()
pybiber_spacy_lg_zscores

pybiber_spacy_trf_zscores <- pybiber_spacy_trf_statistics |>
  get_zscores()
pybiber_spacy_trf_zscores
```

### BiberPlus
```{r}
biberplus_zscores <- biberplus_statistics |>
  mutate(TTR = TTR * 100) |>
  pivot_longer(cols = QUAN:last_col(), names_to = "feature_biberplus") |>
  left_join(feats_loadings |> select(mean, sd, `Linguistic feature`:feature_biberplus, factor1:factor5),
            by = "feature_biberplus") |>
  filter(!is.na(feature_biber)) |>
  mutate(zscore = (value - mean) / sd) |>
  select(doc_id:subcorpus, feature_pseudobiber, value:feature_biber, feature_MAT, feature_biberplus, factor1:zscore) |>
  arrange(corpus, doc_id, subcorpus, feature_pseudobiber)
biberplus_zscores

biberplus_10k_zscores <- biberplus_10k_statistics |>
  mutate(TTR = TTR * 100) |>
  pivot_longer(cols = QUAN:last_col(), names_to = "feature_biberplus") |>
  left_join(feats_loadings |> select(mean, sd, `Linguistic feature`:feature_biberplus, factor1:factor5),
            by = "feature_biberplus") |>
  filter(!is.na(feature_biber)) |>
  mutate(zscore = (value - mean) / sd) |>
  select(doc_id:subcorpus, feature_pseudobiber, value:feature_biber, feature_MAT, feature_biberplus, factor1:zscore) |>
  arrange(corpus, doc_id, subcorpus, feature_pseudobiber)
biberplus_10k_zscores
```

### Biberpy
Note that Biberpy does not extract all features from Biber (1988).
```{r}
biberpy_zscores <- biberpy_statistics |>
  mutate(J43.TTR = J43.TTR * 100) |>
  pivot_longer(cols = A01.pastVerbs:last_col(), names_to = "feature_biberpy") |>
  left_join(feats_loadings |> select(mean, sd, `Linguistic feature`:feature_biberpy, factor1:factor5),
            by = "feature_biberpy") |>
  filter(!is.na(feature_biber)) |>
  mutate(zscore = (value - mean) / sd) |>
  select(doc_id:subcorpus, feature_pseudobiber, value:feature_biber, feature_MAT, feature_biberpy, factor1:zscore) |>
  arrange(corpus, doc_id, subcorpus, feature_pseudobiber)
biberpy_zscores
```




## Computing dimension scores
```{r}
add_zscores <- function(zscores, min_mean = FALSE) {
  zscores <- biber_dims |>
    left_join(zscores, by = "Linguistic feature")
  
  missing_features <- zscores |>
    filter(is.na(doc_id)) |>
    pull(`Linguistic feature`)
  
  if (length(missing_features)) {
    warning("Missing features: ", str_c(missing_features, collapse = ", "))
  }
  
  zscores <- zscores |>
    filter(!is.na(doc_id))
  
  if (min_mean && is.numeric(min_mean) && min_mean > 0) {
    zscores <- zscores |>
      filter(mean > min_mean)
  }
  
  zscores |>
    summarise(Dimension1 = sum(zscore * dim1),
              Dimension2 = sum(zscore * dim2),
              Dimension3 = sum(zscore * dim3),
              Dimension4 = sum(zscore * dim4),
              Dimension5 = sum(zscore * dim5),
              .by = doc_id:subcorpus)
}
```

Adding *z*-scores, not using features with a mean of 1 or below (as proposed by
Nini):
```{r}
min_mean_value <- 1 # to be able to re-run this easily with other values or FALSE

nini_total <- nini_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_stanza_total <- pseudobiber_stanza_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_spacy_sm_total <- pseudobiber_spacy_sm_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_spacy_md_total <- pseudobiber_spacy_md_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_spacy_lg_total <- pseudobiber_spacy_lg_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_sm_total <- pybiber_spacy_sm_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_md_total <- pybiber_spacy_md_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_lg_total <- pybiber_spacy_lg_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_trf_total <- pybiber_spacy_trf_zscores |>
  add_zscores(min_mean = min_mean_value)

biberplus_total <- biberplus_zscores |>
  add_zscores(min_mean = min_mean_value)

biberplus_10k_total <- biberplus_10k_zscores |>
  add_zscores(min_mean = min_mean_value)

biberpy_total <- biberpy_zscores |>
  add_zscores(min_mean = min_mean_value)
```

Same thing, but unrestricted (as in Biber's original study)
```{r}
min_mean_value <- FALSE

nini_total_all <- nini_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_stanza_total_all <- pseudobiber_stanza_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_spacy_sm_total_all <- pseudobiber_spacy_sm_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_spacy_md_total_all <- pseudobiber_spacy_md_zscores |>
  add_zscores(min_mean = min_mean_value)

pseudobiber_spacy_lg_total_all <- pseudobiber_spacy_lg_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_sm_total_all <- pybiber_spacy_sm_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_md_total_all <- pybiber_spacy_md_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_lg_total_all <- pybiber_spacy_lg_zscores |>
  add_zscores(min_mean = min_mean_value)

pybiber_spacy_trf_total_all <- pybiber_spacy_trf_zscores |>
  add_zscores(min_mean = min_mean_value)

biberplus_total_all <- biberplus_zscores |>
  add_zscores(min_mean = min_mean_value)

biberplus_10k_total_all <- biberplus_10k_zscores |>
  add_zscores(min_mean = min_mean_value)

biberpy_total_all <- biberpy_zscores |>
  add_zscores(min_mean = min_mean_value)
```


## Comparisons

### pseudobibeR vs. pybiber (spaCy)
The results from these two should be identical if the same models (sm, md or lg)
are used, so let's compare them first.
```{r}
pb <- bind_rows(
  list(
    pseudobiber_spacy_sm_total |>
      mutate(method = "pseudobibeR w/ spaCy (sm)", .before = doc_id),
    pseudobiber_spacy_md_total |>
      mutate(method = "pseudobibeR w/ spaCy (md)", .before = doc_id),
    pseudobiber_spacy_sm_total |>
      mutate(method = "pseudobibeR w/ spaCy (lg)", .before = doc_id),
    pybiber_spacy_sm_total |>
      mutate(method = "pybiber w/ spaCy (sm)", .before = doc_id),
    pybiber_spacy_md_total |>
      mutate(method = "pybiber w/ spaCy (md)", .before = doc_id),
    pybiber_spacy_sm_total |>
      mutate(method = "pybiber w/ spaCy (lg)", .before = doc_id)
  )
) |>
  mutate(subcorpus = factor(subcorpus,
                            levels = c("quotes", "nonquotes"),
                            labels = c("Dialogue", "Narration")))
```

```{r}
pb |>
  group_by(corpus, subcorpus, method) |>
  skim() |>
  select(-n_missing, -complete_rate)
```


While the values aren't *identical* as they should be (which may be a result of
either pybiber or pseudobibeR not deleting superficial tokens related to UD's
tokenisation preferences), it's hard to spot these differences in the plots --
so neither the choice of pybiber vs. pseudobibeR nor the choice of spaCy's
specific language model should make much of a difference in practice. (But we
haven't looked at the transformer model yet.)
```{r}
#| fig-width: 15
#| fig-height: 15
d1 <- pb |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d2 <- pb |>
  ggplot(aes(x = corpus, y = Dimension2, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d3 <- pb |>
  ggplot(aes(x = corpus, y = Dimension3, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d4 <- pb |>
  ggplot(aes(x = corpus, y = Dimension4, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d5 <- pb |>
  ggplot(aes(x = corpus, y = Dimension5, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d1 + d2 + d3 + d4 + d5 +
  plot_layout(guides = "collect",
              axis_titles = "collect",
              ncol = 1)
```



A closer look at dimension 1, this time without Nini's filter:
```{r}
#| fig-width: 15
pb2 <- bind_rows(
  list(
    pseudobiber_spacy_sm_total_all |>
      mutate(method = "pseudobibeR w/ spaCy (sm)", .before = doc_id),
    pseudobiber_spacy_md_total_all |>
      mutate(method = "pseudobibeR w/ spaCy (md)", .before = doc_id),
    pseudobiber_spacy_sm_total_all |>
      mutate(method = "pseudobibeR w/ spaCy (lg)", .before = doc_id),
    pybiber_spacy_sm_total_all |>
      mutate(method = "pybiber w/ spaCy (sm)", .before = doc_id),
    pybiber_spacy_md_total_all |>
      mutate(method = "pybiber w/ spaCy (md)", .before = doc_id),
    pybiber_spacy_sm_total_all |>
      mutate(method = "pybiber w/ spaCy (lg)", .before = doc_id)
  )
) |>
  mutate(subcorpus = factor(subcorpus,
                            levels = c("quotes", "nonquotes"),
                            labels = c("Dialogue", "Narration")))

pb2 |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), scales = "free_y") +
  scale_colour_paletteer_d("colorblindr::OkabeIto")
```

### Window sizes in BiberPlus
How much does window size influence the results when using BiberPlus?
```{r}
bp <- bind_rows(
  list(
    biberplus_total_all |>
      mutate(method = "BiberPlus (1,000 tokens)", .before = doc_id),
    biberplus_10k_total_all |>
      mutate(method = "BiberPlus (10,000 tokens)", .before = doc_id)
  )
) |>
  mutate(subcorpus = factor(subcorpus,
                            levels = c("quotes", "nonquotes"),
                            labels = c("Dialogue", "Narration")))
```

```{r}
bp |>
  group_by(corpus, subcorpus, method) |>
  skim() |>
  select(-n_missing, -complete_rate)
```

While the differences aren't huge, they are noticeable for some dimensions.
```{r}
#| fig-width: 15
#| fig-height: 15
d1 <- bp |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d2 <- bp |>
  ggplot(aes(x = corpus, y = Dimension2, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d3 <- bp |>
  ggplot(aes(x = corpus, y = Dimension3, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d4 <- bp |>
  ggplot(aes(x = corpus, y = Dimension4, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d5 <- bp |>
  ggplot(aes(x = corpus, y = Dimension5, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method)) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d1 + d2 + d3 + d4 + d5 +
  plot_layout(guides = "collect",
              axis_titles = "collect",
              ncol = 1)
```


### MAT vs. pseudobibeR (Stanza) vs. pybiber (spaCy) vs. BiberPlus vs. Biberpy
Next, we'll compare MAT, pseudobibeR using Stanza, pybibeR using spaCy (only the
small and transformer models for the greatest possible contrast), BiberPlus and
Biberpy.

We'll first include Nini's filter for feature exclusion.
```{r}
full <- bind_rows(
  list(
    nini_total |>
      mutate(method = "MAT", .before = doc_id),
    pseudobiber_stanza_total |>
      mutate(method = "pseudobibeR\n(Stanza)", .before = doc_id),
    pseudobiber_spacy_sm_total |>
      mutate(method = "pseudobibeR\n(spaCy, sm)", .before = doc_id),
    pybiber_spacy_sm_total |>
      mutate(method = "pybiber\n(spaCy, sm)", .before = doc_id),
#    pybiber_spacy_trf_total |>
#      mutate(method = "pybiber\n(spaCy, trf)", .before = doc_id),
    biberplus_total |>
      mutate(method = "BiberPlus\n(spaCy, sm, 1k tokens)", .before = doc_id),
    biberplus_10k_total |>
      mutate(method = "BiberPlus\n(spaCy, sm, 10k tokens)", .before = doc_id),
    biberpy_total |>
      mutate(method = "Biberpy\n(spaCy, sm)", .before = doc_id)
  )
) |>
  mutate(subcorpus = factor(subcorpus,
                            levels = c("quotes", "nonquotes"),
                            labels = c("Dialogue", "Narration")),
         method = fct_inorder(method))

n_comparisons <- full$method |> levels() |> length()
```

Biberpy is the clearest outlier among the different systems.

Results from BiberPlus exhibit the second-largest differences from all others,
especially on dimensions 2 and 5.

Looking at MAT, pseudobibeR and pybiber:

- Interestingly, results from Stanza and spaCy don't seem to differ too much
  from each other (but slightly more than those from different spaCy models).
- MAT's results differ a little more from the others (especially on dimension 3),
  but are surprisingly similar in general.


```{r}
#| fig-width: 12
#| fig-height: 14
d1 <- full |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d2 <- full |>
  ggplot(aes(x = corpus, y = Dimension2, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d3 <- full |>
  ggplot(aes(x = corpus, y = Dimension3, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d4 <- full |>
  ggplot(aes(x = corpus, y = Dimension4, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d5 <- full |>
  ggplot(aes(x = corpus, y = Dimension5, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d1 + d2 + d3 + d4 + d5 +
  plot_layout(guides = "collect",
              axis_titles = "collect",
              ncol = 1)

# only boxplots:
d1 <- full |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 1", colour = "Subcorpus")

d2 <- full |>
  ggplot(aes(x = corpus, y = Dimension2, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 2", colour = "Subcorpus")

d3 <- full |>
  ggplot(aes(x = corpus, y = Dimension3, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 3", colour = "Subcorpus")

d4 <- full |>
  ggplot(aes(x = corpus, y = Dimension4, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 4", colour = "Subcorpus")

d5 <- full |>
  ggplot(aes(x = corpus, y = Dimension5, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 5", colour = "Subcorpus")

# 11.5x11
d2 + d3 + d4 + d5 +
  plot_layout(guides = "collect",
              axis_titles = "collect",
              ncol = 1)
```

If we compare these results to those from Egbert and Mahlberg (2020) -- where
only dimension 1 is available -- we see more pronounced differences. However,
these are likely mostly due to the exclusion of several features in our analysis.
```{r}
#| fig-width: 12
#| fig-height: 3
biber_comparison_filtered <- full_scores_egbert_mahlberg |>
  select(doc_id:subcorpus, Biber_D1) |>
  mutate(method = "Biber Tagger", .before = doc_id,
         subcorpus = factor(subcorpus)) |>
  rename(Dimension1 = Biber_D1) |>
  bind_rows(full |> select(method:Dimension1)) |>
  mutate(method = fct_inorder(method))

biber_comparison_filtered |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons + 1) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

# only boxplots for paper (13x5):
biber_comparison_filtered |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons + 1) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 1", colour = "Subcorpus")
```

If we compare our dimensions using all features associated with them according
to Biber (1988), we get results that are more dissimilar.
```{r}
#| fig-width: 12
#| fig-height: 14
full_all <- bind_rows(
  list(
    nini_total_all |>
      mutate(method = "MAT", .before = doc_id),
    pseudobiber_stanza_total_all |>
      mutate(method = "pseudobibeR\n(Stanza)", .before = doc_id),
    pseudobiber_spacy_sm_total_all |>
      mutate(method = "pseudobibeR\n(spaCy, sm)", .before = doc_id),
    pybiber_spacy_sm_total_all |>
      mutate(method = "pybiber\n(spaCy, sm)", .before = doc_id),
    biberplus_total_all |>
      mutate(method = "BiberPlus\n(spaCy, sm, 1k tokens)", .before = doc_id),
    biberplus_10k_total_all |>
      mutate(method = "BiberPlus\n(spaCy, sm, 10k tokens)", .before = doc_id),
    biberpy_total_all |>
      mutate(method = "Biberpy\n(spaCy, sm)", .before = doc_id)
  )
) |>
  mutate(subcorpus = factor(subcorpus,
                            levels = c("quotes", "nonquotes"),
                            labels = c("Dialogue", "Narration")),
         method = fct_inorder(method))

n_comparisons_all <- full_all$method |> levels() |> length()

d1 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d2 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension2, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d3 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension3, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d4 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension4, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d5 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension5, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

d1 + d2 + d3 + d4 + d5 +
  plot_layout(guides = "collect",
              axis_titles = "collect",
              ncol = 1)

# only boxplots:
d1 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 1", colour = "Subcorpus")

d2 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension2, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 2", colour = "Subcorpus")

d3 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension3, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 3", colour = "Subcorpus")

d4 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension4, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 4", colour = "Subcorpus")

d5 <- full_all |>
  ggplot(aes(x = corpus, y = Dimension5, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons_all) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 5", colour = "Subcorpus")

# 11.5x11
d2 + d3 + d4 + d5 +
  plot_layout(guides = "collect",
              axis_titles = "collect",
              ncol = 1)
```

### Biber Tagger
These are, however, much closer to the results from the Biber Tagger:
```{r}
#| fig-width: 12
#| fig-height: 3
biber_comparison <- full_scores_egbert_mahlberg |>
  select(doc_id:subcorpus, Biber_D1) |>
  mutate(method = "Biber Tagger", .before = doc_id,
         subcorpus = factor(subcorpus)) |>
  rename(Dimension1 = Biber_D1) |>
  bind_rows(full_all |> select(method:Dimension1)) |>
  mutate(method = fct_inorder(method))

biber_comparison |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_violin(position = position_dodge(0.9), aes(fill = after_scale(ggplot2::alpha(colour, 0.1)))) +
  geom_boxplot(position = position_dodge(0.9), width = 0.1, outlier.alpha = .7, outlier.size = 1) +
  facet_wrap(vars(method), ncol = n_comparisons_all + 1) +
  scale_colour_paletteer_d("colorblindr::OkabeIto")

# paper (13x5)
d1_bc <- biber_comparison |>
  ggplot(aes(x = corpus, y = Dimension1, colour = subcorpus)) +
  geom_boxplot(position = position_dodge(0.9)) +
  facet_wrap(vars(method), ncol = n_comparisons_all + 1) +
  scale_colour_paletteer_d("colorblindr::OkabeIto") +
  labs(x = "Corpus", y = "Dimension 1", colour = "Subcorpus")

d1_bc
```



```{r}
full_all <- bind_rows(
  list(
    nini_total_all |>
      mutate(method = "MAT", .before = doc_id),
    pseudobiber_stanza_total_all |>
      mutate(method = "pseudobibeR (Stanza)", .before = doc_id),
    pseudobiber_spacy_sm_total_all |>
      mutate(method = "pseudobibeR (spaCy, sm)", .before = doc_id),
    pseudobiber_spacy_lg_total_all |>
      mutate(method = "pseudobibeR (spaCy, lg)", .before = doc_id),
    pybiber_spacy_sm_total_all |>
      mutate(method = "pybiber (spaCy, sm)", .before = doc_id),
    pybiber_spacy_trf_total_all |>
      mutate(method = "pybiber (spaCy, trf)", .before = doc_id),
    biberplus_total_all |>
      mutate(method = "BiberPlus (spaCy, sm, 1k tokens)", .before = doc_id),
    biberplus_10k_total_all |>
      mutate(method = "BiberPlus (spaCy, sm, 10k tokens)", .before = doc_id),
    biberpy_total_all |>
      mutate(method = "Biberpy (spaCy, sm)", .before = doc_id)
  )
) |>
  mutate(
    subcorpus = factor(subcorpus,
                            levels = c("quotes", "nonquotes"),
                            labels = c("Dialogue", "Narration"))
    )

biber_comparison <- full_scores_egbert_mahlberg |>
  select(doc_id:subcorpus, Biber_D1) |>
  mutate(method = "Biber Tagger", .before = doc_id,
         subcorpus = factor(subcorpus)) |>
  rename(Dimension1 = Biber_D1) |>
  bind_rows(full_all |> select(method:Dimension1)) |>
  mutate(method = fct_inorder(method))


mixed_mod <- lmer(Dimension1 ~ corpus + subcorpus + corpus:subcorpus + method + (1|doc_id),
                  biber_comparison)

model_parameters(mixed_mod)

model_parameters(mixed_mod, standardize = "refit")

model_performance(mixed_mod)
```

```{r}
#| fig-width: 18
#| fig-height: 10
check_model(mixed_mod)
```

Stargazer output for the paper:
```{r}
#| warning: false
stargazer(mixed_mod, type = "text",
          ci = TRUE, digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001))
```


As requested by one reviewer, here are scatter plots to show the dimension 1
scores of individual documents -- always comparing the Biber Tagger against one
other system. The dashed line represents perfect correlation. Points under the
line mean lower dimension 1 scores for the system in question than for the Biber
Tagger, points over the line mean higher scores.
```{r}
biber_comparison_1v1 <- biber_comparison |>
  pivot_wider(names_from = method, values_from = Dimension1)

plot_1v1 <- function(method) {
  biber_comparison_1v1 |>
    ggplot(aes(x = `Biber Tagger`, y = !!as.name(method), colour = subcorpus, shape = corpus, label = doc_id)) +
    geom_point() +
    geom_text(size = 2, vjust = 1.2, colour = "darkgrey") +
    geom_abline(alpha = .5, linetype = 2)
}

systems <- levels(biber_comparison$method)
systems <- systems[systems != "Biber Tagger"]

for (s in systems) {
  print(plot_1v1(s))
  Sys.sleep(.5)
}
```



## Feature-level comparisons
To find out where the differences between individual systems/language models are
coming from, we can look at the features underlying the register dimensions.

For this comparison, we'll have a closer look at MAT, pseudobibeR using Stanza
and spaCy (large language model), pybiber (using the same spaCy model) and
BiberPlus.
```{r}
mat_texts <- nini_zscores |>
  select(doc_id:subcorpus, `Linguistic feature`, feature_pseudobiber, mean, value, zscore) |>
  arrange(doc_id, corpus, subcorpus, feature_pseudobiber) |>
  mutate(system = "MAT: Stanford tagger", .before = "doc_id")

pseudobiber_texts_stanza <- pseudobiber_stanza_zscores |>
  select(doc_id:subcorpus, `Linguistic feature`, feature_pseudobiber, mean, value, zscore) |>
  arrange(doc_id, corpus, subcorpus, feature_pseudobiber) |>
  mutate(system = "pseudobibeR: Stanza", .before = "doc_id")

pseudobiber_texts_spacy_sm <- pseudobiber_spacy_sm_zscores |>
  select(doc_id:subcorpus, `Linguistic feature`, feature_pseudobiber, mean, value, zscore) |>
  arrange(doc_id, corpus, subcorpus, feature_pseudobiber) |>
  mutate(system = "pseudobibeR: spaCy (sm)", .before = "doc_id")

pybiber_texts_spacy_sm <- pybiber_spacy_sm_zscores |>
  select(doc_id:subcorpus, `Linguistic feature`, feature_pseudobiber, mean, value, zscore) |>
  arrange(doc_id, corpus, subcorpus, feature_pseudobiber) |>
  mutate(system = "pybiber: spaCy (sm)", .before = "doc_id")

biberplus_texts <- biberplus_zscores |>
  select(doc_id:subcorpus, `Linguistic feature`, feature_pseudobiber, mean, value, zscore) |>
  arrange(doc_id, corpus, subcorpus, feature_pseudobiber) |>
  mutate(system = "BiberPlus: spaCy (sm)", .before = "doc_id")

biberpy_texts <- biberpy_zscores |>
  select(doc_id:subcorpus, `Linguistic feature`, feature_pseudobiber, mean, value, zscore) |>
  arrange(doc_id, corpus, subcorpus, feature_pseudobiber) |>
  mutate(system = "Biberpy: spaCy (sm)", .before = "doc_id")

comparison <- bind_rows(mat_texts,
                        pseudobiber_texts_stanza,
                        pseudobiber_texts_spacy_sm,
                        pybiber_texts_spacy_sm,
                        biberplus_texts,
                        biberpy_texts) |>
  inner_join(biber_dims, by = "Linguistic feature") |>
  arrange(doc_id, corpus, subcorpus, feature_pseudobiber, system) |>
  mutate(system = factor(system,
                         levels = c("MAT: Stanford tagger", "pseudobibeR: Stanza", "pseudobibeR: spaCy (sm)", "pybiber: spaCy (sm)", "BiberPlus: spaCy (sm)", "Biberpy: spaCy (sm)")))

# comparison |> write_csv("comparison_mat_pseudobiber_pybiber_biberplus_biberpy.csv")
```

### Dimension 1
Without filtering, we can immediately see that feature #23 (WH questions) takes
suspiciously high values for Biberpy. This makes it a little harder to see
details about the other features.

```{r}
#| fig-width: 18
#| fig-height: 10
comparison |>
  filter(dim1 != 0) |>
  mutate(contribution = ifelse(dim1 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 1")
```

We'll use an upper limit for the y-axis to see things a little more clearly.


Features that would be removed by the filter:

- 13 (WH questions) is suspiciously high in Biberpy
- 23 (WH clauses) is lower for pybiber than for all others; the count is very
  often 0 for individual documents, as pybiber subtracts the counts of features
  31 and 32
- 34 (sentence relatives) is practically identical for all systems except
  Biberpy (much higher)
- 47 (hedges) is slightly lower for MAT and Biberpy


Other observations:

- pseudobibeR, whether with spaCy or Stanza, and pybiber agree on attributive
  adjectives, as do MAT, BiberPlus and Biberpy -- at a significantly higher
  level
- BiberPlus and Biberpy also agree on nouns, but all other systems count more;
  something similar can be observed for word length and type-token ratio (which
  shouldn't be all that controversial)
- BiberPlus counts significantly more prepositions than the other systems
- Biberpy is the lone outlier for demonstrative pronouns, discourse particles,
  emphatics and sentence relatives; it also has very high values for THAT
  deletions
- MAT and BiberPlus count more non-phrasal coordinations
- pybiber changed its computation of BE as main verb (with an option to revert
  it to pseudobibeR's behaviour); it now counts only counts it if a form of *to
  be* is the root of the sentence (whereas Biber's original rule is more lax);
  this results in substantially lower counts
```{r}
#| fig-width: 18
#| fig-height: 10
dim1_comp <- comparison |>
  filter(dim1 != 0) |>
  mutate(contribution = ifelse(dim1 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + # linewidth = .4, outlier.size = 1
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 1") +
  coord_cartesian(ylim=c(NA, 10))
dim1_comp
```


Only keeping features with a mean above 1 in Biber (1988) leaves us with 24 out
of 28 features:
```{r}
#| fig-width: 18
#| fig-height: 10
comparison |>
  filter(dim1 != 0,
         mean > 1) |>
  mutate(contribution = ifelse(dim1 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 1")
```


### Dimension 2
BiberPlus has slightly lower values for all features than MAT, pseudobibeR and
pybiber. Biberpy is missing two features, and counts for past tense and
synthetic negatives are even lower than those from BiberPlus.

Without filtering:
```{r}
dim2_comp <- comparison |>
  filter(dim2 != 0) |>
  mutate(contribution = ifelse(dim2 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + # linewidth = .4, outlier.size = 1
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 2")
dim2_comp
```

With filtering, 5 out of 6 features are used: #25 (present participial clauses)
not included -- same for all taggers except Biberpy:
```{r}
comparison |>
  filter(dim2 != 0,
         mean > 1) |>
  mutate(contribution = ifelse(dim2 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 2")
```


### Dimension 3
Most obvious differences:

- much lower count of adverbs for MAT (and Biberpy, to a lesser extent) → much
  higher dim. 3
- slightly higher count of time adverbials for MAT → lower dim. 3
- lower count of WH relatives in subject position for MAT → lower dim. 3
- higher count of WH relatives in object position for BiberPlus (somewhat offset
  by slightly lower counts for WH relatives in subject position and phrasal
  coordinations, but a slightly lower adverb count also adds to the dim. 3
  scores)

Without filtering:
```{r}
dim3_comp <- comparison |>
  filter(dim3 != 0) |>
  mutate(contribution = ifelse(dim3 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + # linewidth = .4, outlier.size = 1
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 3")
dim3_comp
```

With filtering, 7 out of 8 features are used: 33 (WH relatives: pied pipes) not
included -- similar for all taggers (slightly lower for Biberpy)
```{r}
comparison |>
  filter(dim3 != 0,
         mean > 1) |> # 
  mutate(contribution = ifelse(dim3 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 3")
```



### Dimension 4
Since all feature means for dimension 4 in the original study are above 1, all
features are included with or without filtering. Differences between MAT,
pseudobibeR and pybiber are marginal -- MAT only counts a few more split
auxiliaries. BiberPlus, however, counts a lot more split auxiliaries which would
result in much higher dimension 4 values if not for slightly lower counts for
four of the other features.
```{r}
dim4_comp <- comparison |>
  filter(dim4 != 0) |> # uses 6/6 features (filter makes no difference)
  mutate(contribution = ifelse(dim4 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + # linewidth = .4, outlier.size = 1
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 4")
dim4_comp
```


### Dimension 5
It's hard to see much in this plot, as Biberpy's count of conjuncts seems to be
absurdly high (also very high for BiberPlus).
```{r}
comparison |>
  filter(dim5 != 0) |>
  mutate(contribution = ifelse(dim5 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + # linewidth = .4, outlier.size = 1
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 5")
```

For the plot's sake, let's remove Biberpy's count of conjuncts from the picture:
```{r}
dim5_comp <- comparison |>
  filter(dim5 != 0) |>
  filter(!(feature_pseudobiber == "f_45_conjuncts" & system == "Biberpy: spaCy (sm)")) |>
  mutate(contribution = ifelse(dim5 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) + # linewidth = .4, outlier.size = 1
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 5")
dim5_comp
```


With filtering, only 3 out of 6 features are used. The following are discarded:

- 18 (BY passives)
- 26 (past participial clauses)
- 38 (adv. subordinator - other)

Feature 38 in particular is much lower for MAT than for the other systems. This
alone can explain the lower scores on dimension 5 (as well as the higher scores
for Stanza).
```{r}
comparison |>
  filter(dim5 != 0,
         mean > 1) |>
  mutate(contribution = ifelse(dim5 > 0, "positive", "negative")) |>
  ggplot(aes(x = `Linguistic feature`, y = zscore, colour = system)) +
  geom_boxplot(outlier.size = 1, outlier.alpha = .5) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0), axis.title.y = element_markdown()) +
  facet_grid(cols = vars(contribution), scales = "free_x", space = "free") +
  scale_colour_oi(order = c(1, 3, 2, 5, 7, 8, 4, 6, 9)) +
  labs(y = "<i>z</i>-score", colour = "System", title = "Dimension 5")
```


```{r}
#| fig-width: 18
#| fig-height: 10
design <- "
  11
  23
  45
"

dim1_comp + dim2_comp + dim3_comp + dim4_comp + dim5_comp +
  plot_layout(guides = "collect",
              axis_titles = "collect",
              design = design)
# 11x11
```


## Speed comparisons
Speed comparisons favour pybiber over pseudobibeR. All Python systems using
spaCy are almost equally fast. Using the GPU instead of the CPU speeds spaCy up
considerably.

spaCy's transformer model (which requires GPU usage) is a lot slower than its
more lightweight models. Similarly, Stanza itself, which is also
transformer-based, takes slightly over an hour to annotate the whole corpus.
```{r}
#| message: false
speed_pseudobiber <- read_csv("output_pseudobiber/quotes_nonquotes_corpus/spacy_timings_ttr100.csv") |>
  mutate(program = "pseudobibeR", .before = model,
         time_minutes = round(time_sec / 60)) |>
  select(-time_sec)

speed_pybiber <- read_csv("output_pybiber/timing_results.csv") |>
  mutate(program = "pybiber", .before = model,
         time_minutes = round(time_seconds / 60, 2)) |>
  select(-time_seconds)

speed_biberplus <- read_csv("output_biberplus/timing_results_no_gpu.csv") |>
  mutate(program = "BiberPlus", .before = model,
         time_minutes = round(time_seconds / 60, 2)) |>
  select(-time_seconds)

speed_biberplus_gpu <- read_csv("output_biberplus/timing_results.csv") |>
  mutate(program = "BiberPlus (GPU)", .before = model,
         time_minutes = round(time_seconds / 60, 2)) |>
  select(-time_seconds)

speed_biberpy <- read_csv("output_biberpy/timing_results.csv") |>
  mutate(program = "Biberpy", .before = model,
         model = str_remove(model, "en_core_web_"),
         time_minutes = round((time_seconds + 67) / 60, 2)) |> # add runtime of biber-dym.py
  select(-time_seconds)

speed_stanza <- read_csv("output_stanza/timing_results.csv") |>
  mutate(program = "Stanza", .before = model,
         model = "combined model",
         time_minutes = round(time_seconds / 60, 2)) |>
  select(-time_seconds)

speeds <- bind_rows(speed_pseudobiber, speed_pybiber, speed_biberplus, speed_biberplus_gpu, speed_biberpy, speed_stanza)
speeds
```